---
title: "IODS Final assignment"
author: "Tuomas Tiihonen"
date: "10 joulukuuta 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Preparatory actions

Before we can start analysing the student data, some groundwork is needed to prepare the environment for the analysis, and actually import the data as well.

Much of the data preparation work towards better analysis possibilities with MCA has been done already in the [data wrangling script](https://github.com/fanders-iods/IODS-final/blob/master/data/wrangle_students.R).

## Library import

We'll start off by importing the necessary libraries that will provide us the many of the functions we'll use in this assignment. Because the libraries merely provide us the functions we'll describe when they're used in the context of the data, I've included short descriptions of the libraries as code comments below.

```{r library access, echo=TRUE}
library(FactoMineR) # This includes a lot of factor analysis stuff
library(factoextra) # Considered, if you will, an extension to FactoMiner with more progressive plotting capabilities
library(ggplot2) # The do-all plotting library
library(dplyr) # Everyone needs to manipulate data
library(corrplot) # Here to provide more beautiful correlation coefficient plotting
library(tidyr) # Tidyr provides us data manipulation functions
```

## Data import

The next bit is importing the data and taking a look at it.

```{r data import}
load(file = "/Volumes/tuti/IODS-final/data/wrangled_students.Rdata")
dim(rawdata)
```

The data has 382 observations of 38 variables. Let's next look at the variables.

```{r}
str(rawdata)
```

Further information on all of the original variables, as well as the origin of the data, is available from the [source](https://archive.ics.uci.edu/ml/datasets/STUDENT+ALCOHOL+CONSUMPTION) 

A short recap of the variables in the wrangled dataset:

* school: student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira) 
* sex: student's sex (binary: 'F' - female or 'M' - male) 
* age: student's age (numeric: from 15 to 22) 
* address: student's home address type (binary: 'U' - urban or 'R' - rural)
* famsize: family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3) 
* Pstatus: parent's cohabitation status (binary: 'T' - living together or 'A' - apart) 
* Medu: mother's education (binary: "high" = secondary or higher, "low" = lower than secondary)
* Fedu: father's education (binary: "high" = secondary or higher, "low" = lower than secondary)
* Mjob: mother's job (factor: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
* Fjob: father's job (factor: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other') 
* reason: reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other') 
* nursery: attended nursery school (binary: yes or no) 
* internet: Internet access at home (binary: yes or no) 
* guardian: student's guardian (factor: 'mother', 'father' or 'other')
* traveltime: home to school travel time (factor: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour) 
* studytime: weekly study time (factor: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours) 
* failures: past class failures (binary: true or false)
* schoolsup: extra educational support (binary: yes or no) 
* famsup: family educational support (binary: yes or no) 
* paid: extra paid classes within the course subject (Math or Portuguese) (binary: yes or no) 
* activities: extra-curricular activities (binary: yes or no)
* higher: wants to take higher education (binary: yes or no) 
* romantic: with a romantic relationship (binary: yes or no) 
* famrel: quality of family relationships (numeric: from 1 - very bad to 5 - excellent) 
* freetime: free time after school (numeric: from 1 - very low to 5 - very high) 
* goout: going out with friends (numeric: from 1 - very low to 5 - very high) 
* health: current health status (numeric: from 1 - very bad to 5 - very good)
* absences: The student's absences.
* G3: final grade (numeric: from 0 to 20)
* high_use: does the student use high amount of alcohol (binary: true or false)
* G3_quart: The quartile of the student's final grade based on G3.

Descriptions for unmodified variables are from the [source](https://archive.ics.uci.edu/ml/datasets/STUDENT+ALCOHOL+CONSUMPTION) mutatis mutandis factorised variables. Variables created during data wrangling are described by the author.

Finally, let's see how the observations are divided in the data. We'll use both the classical summary function and a graphical overview. As we've got lots of variables, we also define a higher width and height for the graphics to display nicely.

## Exploring the variables

It pays to take a graphical look at the variables. People may have different ways of making sense of data containing multiple variables, but a graphical representation usually works at least for finding out whether there are some variables that are so skewed that they would also skew any further analysis.

```{r, fig.width = 12, fig.height = 12}
summary(rawdata)

gather(rawdata) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

From these overviews we can see deduce the following things that are pertinent for further analysis:

* Only a small number of students have guardians outside their parents. On the other hand, co-guardianship has not been measured separately.
* Only a small number of fathers either stay at home or work in healthcare.
* When taken together with the mothers' employment, the number of parents employed as teachers is ~12%, which is quite well above the OECD average of 3.94%, or the national front runner Iceland with 7.8% ([data](http://www.nationmaster.com/country-info/stats/Education/Teachers-as-percentage-of-labor-force) from 1999). Hence further exploration of the participating schools would be warranted to gauge their representativeness of the Portuguese students' backgrounds.
* A negligible number of students do not strive for higher education.
* The number of students without Internet access from home is small, but non-negligible.

# Hypotheses

The null hypothesis is that no significant categories can be established in the data. The alternate hypothesis is that the education levels of the parents are categorised with a high quartile of the final grade.